---
title: "rmarkdown_1"
author: "Yinan Fneg"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load needed packages

```{r}
library(tidyverse)
library(sf)
library(ggspatial)
library(ggplot2)
```
## Load the data 

Load the ‘knox_acs.shp’ shapefile into R and give it the name ‘knox’.


```{r echo=FALSE}
knox <- st_read('./doc/lab02_canvas/lab02_canvas/knox_acs.shp')
```
## Add column

Add a new column to knox that calculates the fraction of each census tract that is between ages 18 and 21. Call that column “college_frac”.

```{r echo=FALSE}
knox <- knox %>%
  mutate(college_frac = AGE18_21 / TOTAL_POP)
```

## Choropleth plot

Create a choropleth map of the proportion of the population age 18-21. Make sure that the map is appropriately labeled.

```{r echo=FALSE}
ggplot(knox) + 
  annotation_map_tile() +
  geom_sf(aes(fill=college_frac)) + 
  scale_fill_gradient(name = "Proportion 18-21") +
  labs(title = "Proportion of Population Aged 18-21 by Census Tract") +
  coord_sf(datum=NA)
```

## Histogram plot

Plot a histogram of the Median Household Income and label the axes appropriately. In words, describe the shape of this distribution.

```{r echo=FALSE}
ggplot(knox, aes(x = MEDHHINC/1000)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Median Household Income \n($1000s)",
       x = "Median Household Income",
       y = "Number of Census Tracts \n($1000s)") +
  theme_minimal()
```