---
title: "rmarkdown_1"
author: "Yinan Fneg"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load needed packages

```{r}
library(tidyverse)
library(sf)
library(ggspatial)
library(ggplot2)
```
## Load the data 

Load the ‘knox_acs.shp’ shapefile into R and give it the name ‘knox’.


```{r echo=FALSE}
knox <- st_read('./doc/lab02_canvas/lab02_canvas/knox_acs.shp')
```
## Add column

Add a new column to knox that calculates the fraction of each census tract that is between ages 18 and 21. Call that column “college_frac”.

```{r echo=FALSE}
knox <- knox %>%
  mutate(college_frac = AGE18_21 / TOTAL_POP)
```

## Choropleth plot

Create a choropleth map of the proportion of the population age 18-21. Make sure that the map is appropriately labeled.

```{r echo=FALSE}
ggplot(knox) + 
  annotation_map_tile() +
  geom_sf(aes(fill=college_frac)) + 
  scale_fill_gradient(name = "Proportion 18-21") +
  labs(title = "Proportion of Population Aged 18-21 by Census Tract") +
  coord_sf(datum=NA)
```

## Histogram plot

Plot a histogram of the Median Household Income and label the axes appropriately. In words, describe the shape of this distribution.

```{r echo=FALSE}
ggplot(knox, aes(x = MEDHHINC/1000)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Median Household Income \n($1000s)",
       x = "Median Household Income",
       y = "Number of Census Tracts \n($1000s)")
```

## Histogram plot

Create a scatterplot with Household Income (in $1000s) on the x-axis and House Value (in $1000s) on the y-axis.  Set the alpha level to the population of each tract. Describe in words the shape of this scatterplot.

```{r echo=FALSE}

ggplot(knox, aes(x = MEDHHINC / 1000, y = MEDHVALUE / 1000, alpha = TOTAL_POP)) +
  geom_point(aes(size = TOTAL_POP), color = "black") +
  labs(title = "Scatterplot of Household Income vs. House Value",
       x = "Household Income (in $1000s)",
       y = "House Value (in $1000s)",
       alpha = "Population") +
  theme_minimal() +
  guides(size = FALSE)  # To not show legend for size

```

## Correlation

Calculate the three correlation coefficients between Household Income, House Value, and percent 18-21. Use these correlation coefficients in a sentence without typing the numbers in directly. (i.e. your markdown file should generate the coefficient automatically for you).

```{r echo=FALSE}

summary_stat <- knox %>%
  summarize(
    mean_HHINC = mean(MEDHHINC, na.rm=TRUE),
    sd_HHINC = sd(MEDHHINC, na.rm=TRUE),
    mean_HVALUE = mean(MEDHVALUE, na.rm=TRUE),
    sd_HVALUE = sd(MEDHVALUE, na.rm=TRUE),
    cor_HVALUE_HHINC = cor(MEDHHINC, MEDHVALUE, use='complete.obs'),
    cor_HHINC_1821 = cor(MEDHHINC, college_frac, use='complete.obs'),
    cor_HVALUE_1821 = cor(MEDHVALUE, college_frac, use='complete.obs')
  )

```

The correlation coefficient between Household Income (`MEDHHINC`) and House Value (`MEDHVALUE`) is `r round(summary_stat$cor_HVALUE_HHINC, 2)`. 
The correlation coefficient between Household Income (`MEDHHINC`) and the percentage of the population aged 18-21 (`college_frac`) is `r round(summary_stat$cor_HHINC_1821, 2)`.
Lastly, the correlation coefficient between House Value (`MEDHVALUE`) and the percentage of the population aged 18-21 (`college_frac`) is `r round(summary_stat$cor_HVALUE_1821, 2)`.